Hierarchical wrinkling in a confined permeable biogel

### 2015-01-26 fig_plis/article_plis.tex
Justifie rapidement la forme fractale

\section*{Methods summary}
To prevent adsorption of caseins to surfaces, microscope slides and cover slips were cleaned and subsequently coated with a silane agent (3-(trimethoxysilyl) propyl methacrylate, Sigma Aldrich) to which a polyacrylamide brush was polymerized from methacrylate groups.
Sodium caseinate (Firmenich) was labelled with Dylight~550 NHS ESTER (Thermo Scientific). Excess dye was removed by centrifugation. Unlabelled sodium caseinate powder was dissolved in deionised water (or water--glycerol mixture). To induce gelation glucono-$\delta$-lactone (GDL) in powder (Firmenich) was added to the solution. \SI{45}{\micro\litre} of this solution was immediately mixed with \SI{5}{\micro\litre} of labelled caseinate solution, stirred and injected in the microscopy cell, immediately sealed using ultraviolet-cured glue (Norland Optical) and placed under the microscope before gelation could take place. 
3D data were collected on a Zeiss LSM510 confocal microscope, using \SI{532}{\nano\meter} laser excitation. We used either a 10x (air) lens to measure the geometry of wrinkling (see Supplementary Material) or a 63x (oil) lens to observe the microstructure of the gel. Larger scope pictures, obtained either by stitching fluorescent microscopy images (Nikon Eclipse Ti) or by transmitted/reflected light macroscope (Nikon SMZ745T/Leica DMS1000) were used to measure the wavelength of the primary pattern (Supplementary Figures~\ref{fig:patterns} and \ref{fig:generations}). Secondary pattern wavelength is set to half of the primary wavelength, according to Supplementary Fig.~\ref{fig:generations}.
Data shown in Fig.~\ref{fig:acidification} a-b were obtained with a SevenCompact pH-meter (Mettler Toledo) and an MCR 301 rheometer (Anton Paar).
Permeability is measured using a protocol found in \cite{VanDijk1986} (see Materials and Supplementary Fig.~\ref{fig:permeability}).

@@

\begin{figure}[b]
	\includegraphics{nonmonotonic}%
	\caption{Microstructure evolution. (a) Details of confocal images in fully adhesive conditions (top row) and with no adhesion on top (bottom row). The scale bar is \SI{10}{\micro\metre}. Arrows indicates the pore size. (b) Corresponding Fourier spectra in second situation and same times. Lines are fits by Eq.~\ref{eq:fractalS}. (c-d) Evolution of susceptibility and fractal dimension in both situations (black line and orange dashed line respectively). Symbols indicates the times of the images in (a).}%
	\label{fig:nonmonotonic}%
\end{figure}

To follow the evolution of the microstructure of the gel, we acquired slices of $(\SI{2048}{px})^2$ with pixels of \SI{98}{\nano\metre} (Fig.~\ref{fig:nonmonotonic}a). After Hann windowing we performed Fourier transform and radially averaged the resulting spectrum to obtain $I(q)$ (see Fig.~\ref{fig:nonmonotonic}b) that was fitted by the following fractal form:
\begin{equation}
I(q) = \frac{2\chi\Gamma(d)}{\left(1+\left((d+1)\xi q\right)^2\right)^{d/2}}
\label{eq:fractalS}
\end{equation}
where the susceptibility $\chi$ is related to the fluorescence contrast between the network and the solvent, the cut off length $\xi$ corresponds to the largest pore size and $d$ is the fractal dimension. $\Gamma$ is the gamma function.

>>>>

\section*{Methods summary}
To prevent adsorption of caseins to surfaces, microscope slides and cover slips were cleaned and subsequently coated with a silane agent (3-(trimethoxysilyl) propyl methacrylate, Sigma Aldrich) to which a polyacrylamide brush was polymerized from methacrylate groups.
Sodium caseinate (Firmenich) was labelled with Dylight~550 NHS ESTER (Thermo Scientific). Excess dye was removed by centrifugation. Unlabelled sodium caseinate powder was dissolved in deionised water (or water--glycerol mixture). To induce gelation glucono-$\delta$-lactone (GDL) in powder (Firmenich) was added to the solution. \SI{45}{\micro\litre} of this solution was immediately mixed with \SI{5}{\micro\litre} of labelled caseinate solution, stirred and injected in the microscopy cell, immediately sealed using ultraviolet-cured glue (Norland Optical) and placed under the microscope before gelation could take place. 
3D data were collected on a Zeiss LSM510 confocal microscope, using \SI{532}{\nano\meter} laser excitation. We used either a 10x (air) lens to measure the geometry of wrinkling (see Supplementary Material) or a 63x (oil) lens to observe the microstructure of the gel. Larger scope pictures, obtained either by stitching fluorescent microscopy images (Nikon Eclipse Ti) or by transmitted/reflected light macroscope (Nikon SMZ745T/Leica DMS1000) were used to measure the wavelength of the primary pattern (Supplementary Figures~\ref{fig:patterns} and \ref{fig:generations}). Secondary pattern wavelength is set to half of the primary wavelength, according to Supplementary Fig.~\ref{fig:generations}.
Data shown in Fig.~\ref{fig:acidification} a-b were obtained with a SevenCompact pH-meter (Mettler Toledo) and an MCR 301 rheometer (Anton Paar).
Permeability is measured using a protocol found in \cite{VanDijk1986} (see Supplementary Method and Supplementary Fig.~\ref{fig:permeability}).

@@

\begin{figure}[b]
	\includegraphics{nonmonotonic}%
	\caption{Microstructure evolution. (a) Details of confocal images in fully adhesive conditions (top row) and with no adhesion on top (bottom row). The scale bar is \SI{10}{\micro\metre}. Arrows indicates the pore size. (b) Corresponding Fourier spectra when no adhesion on top and same times. Lines are fits by Eq.~\ref{eq:fractalS}. (c-d) Evolution of susceptibility and fractal dimension in full adhesive (black line) and no adhesion on top (orange dashed line) situations. Symbols indicates the times of the images in (a).}%
	\label{fig:nonmonotonic}%
\end{figure}

To follow the evolution of the microstructure of the gel, we acquired slices of $(\SI{2048}{px})^2$ with pixels of \SI{98}{\nano\metre} (Fig.~\ref{fig:nonmonotonic}a). After Hann windowing we performed Fourier transform and radially averaged the resulting spectrum to obtain $I(q)$ (see Fig.~\ref{fig:nonmonotonic}b) that was fitted by the following fractal form:
\begin{equation}
I(q) = \frac{2\chi\Gamma(d)}{\left(1+\left((d+1)\xi q\right)^2\right)^{d/2}}
\label{eq:fractalS}
\end{equation}
where the susceptibility $\chi$ is related to the fluorescence contrast between the network and the solvent, the cut off length $\xi$ corresponds to the largest pore size and $d$ is the fractal dimension. $\Gamma$ is the gamma function. Eq.~(\ref{eq:fractalS}) is a good approximation to the 2D Fourier transform of a fractal pair correlation function $g(r) \sim 1+ e^{-r/\xi} r^{d-2}$.
